---
title: "Covid91 vaccine study Final2021"
author: "Bessie Luevano"
date: "Summer 2021"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData$treatment <- as.factor(FinalData$treatment)
FinalData$sex <- as.factor(FinalData$sex)
FinalData$infected <- as.factor(FinalData$infected)
FinalData$LGBTQ <- as.factor(FinalData$LGBTQ)
FinalData$DrugUser <- as.factor(FinalData$DrugUser)
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods

The first step is to isolate the subpopulations
```{r}
maledata<- filter(FinalData,sex=="guy")
maledata$sex <- as.factor(maledata$sex)
```


# Guys "sex = guy"

## Bar Chart 

```{r}
Guys <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Guys,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Guys,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

It appears in the bar chart that around 58% of males that were given the drug got Covid-19. About 42% of males that got Covid-19 were given the placebo. It appears that the bar chart shows that males who received the vaccine were about 50% more likely to contract covid-19. 

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The numerical summary shows that 337 males got infected with Covid-19 out of a total of 10,450 males in the study. It appears in the table that 3.72% of the males that received the drug got covid-19 and 2.55% of the males that received the placebo got covid-19. It doesn't appear that the drug is not effective at preventing covid within this study of males.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the table using Chi-squared test shows that the p-vaule is 0.0006181 which is smaller that 1. It appears in the Fisher's Exact Test that you are 1.47% more likely to contract covid-19 if you take the drug. The Effectiveness of the drug is 59.5% using the formula 1.47/2.47*100= 59.5%. Therefore the vaccine is only 40% effective in th male population and doesn't seem to be effective. We can reject the null hypothesis because the drug is ineffective in preventing Covid-19 within males. 

# Gals "sex = gal"

```{r}
femaledata<- filter(FinalData,sex=="gal")
femaledata$sex <- as.factor(femaledata$sex)
```

## Bar Chart 

```{r}
Gals <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Gals,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Gals,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

It appears from the bar chart that about 33% of females that took the drug contracted covid-19. About 66% of females that got the placebo contracted covid-19. It appears from the chart that less than 50% of females who took the drug contracted covid-19.

## Numerical Summary

```{r}
table2 <- xtabs(~treatment + infected, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

Looking at the numerical summary it shows that 607 femaleswere infected with covid-19 from a study of 8,606 females. It appears in the table that 4.44% of females that received the drug got covid-19 and 8.67% of the females that received the placebo contracted covid-19. It appears that the drug is some what effective in preventing covid-19 within this study of females. 

## Inferential

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

In the table Chi-Square test results it shows that the p-vaule is 3.626e-16. Using Fisher's exact test the odds ratio shows that you are 0.489% likely to contract covid-19 if you take the drug. The vaccine is 67% effective in females using 2.04/(1+2.045)*100

# Sexual Orientation "LGBTQ = gay"

```{r}
LGBTQdata<- filter(FinalData,LGBTQ=="gay")
LGBTQdata$LGBTQ <- as.factor(LGBTQdata$LGBTQ)
```

## Bar Chart 

```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


# Drug Users "Drug User = yes"

```{r}
Druggiedata<- filter(FinalData,DrugUser=="yes")
Druggiedata$DrugUser <- as.factor(Druggiedata$DrugUser)
```

## Bar Chart 

```{r}
Druggie <- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggie,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggie,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Using the bar chart for the study of Drug Users, about 34% of those that received the drug contracted covid-19 and about 66% of those that received the placebo contracted covid-19. It appears that those drug users that received the vaccine, less that 50% contracted covid-19.

## Numerical Summary

```{r}
table4 <- xtabs(~treatment + infected, data=DrugUserdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

The numerical summary shows that 117 drug users were infected with covid-19 out of 2,166 drug users in this study. It appers in the table that 3.63% of drug users that received the drug contracted covid-19 and 6.46% that received the placebo contracted covid-19. It appears that the drug is effective in preventing covid-19 in drug users within this study.

## Inferential

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The effectiveness of the vaccine is 35%.



# Discussion & Conclusion










